Formalizing Stack Safety as a Security Property

### 2021-07-09 submission3/main.tex
Under limit!

That said, our model is meant to apply generically to any enforcement mechanism,
including those that do not take advantage of this separation.
Some mechanisms, like hardware capability systems, might exhibit failstop behavior,
but have no protected policy state---this is represented by a trivial policy
that failstops when the hardware faults. Others might never even failstop, such as
code rewriting techniques that turn stack safety violations into other (erroneous, but not
stack-unsafe) behaviors. Then the policy-enhanced machine is identical to the original.

@@

We now define our machine and enforcement model precisely. To make our definition of stack safety
as generic as possible, this section proposes an abstract interface to a
machine model plus some kind of (software- or hardware-enforced) policy monitor.
In~\cref{sec:enforcement}, we instantiate this model with a concrete dynamic
policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.

@@

\paragraphx{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of stack safety
%begins with \citet{SkorstengaardLocal}, who introduced a calling
%convention that uses local capabilities to preserve local state
%encapsulation and well-bracketed control flow
%using a logical relation to reason about the stack safety of concrete
%programs.
% Although the required hardware support is readily available
%in capability machines like Cheri,
%% this technique incurs significant
%% costs, because
%it requires the entire unused part of the stack to be
%cleared whenever a security boundary is crossed.
%Their logical
%relation captures capability safety without ``externally observable
%side-effects (like console output or memory access traces)''
%and can be used to reason about individual programs.
is that of \citet{SkorstengaardSTKJFP}. Their
StkTokens system enforces stack safety on hardware equipped with
memory capabilities, similar to those used in the Cheri system~\citep{Woodruff+14,Chisnall+15}, but
with additional support for linear capabilities.
Stack safety is defined as the conjuncton of local state encapsulation (LSE) and
well-bracketed control flow (WBCF).
Informally, LSE means that the current activiation can only access its own frame,
and WBCF means that the topmost frame can only return to
the immediately adjacent frame below. Formally, these properties are defined
with reference to the semantics of a
capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine.  They call this approach a ``fully abstract overlay semantics.''
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
\begin{enumerate}[label=(\roman*)]
\item Our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
  \citet{SkorstengaardSTKJFP} describe a class of ``unreasonable'' code components
  that are not protected, because the are capable of dynamic behavior that can compromise
  their own security. Unreasonable code can be viewed as adversarial, but it is equally
  likely to simply be buggy. One can imagine a function that, when called with the wrong
  inputs, behaves unreasonably, but under most circumstances can be
  considered reasonable. \citet{SkorstengaardSTKJFP} would not consider this a
  {\em reasonable component} (as defined in their section 4.3.3), and so
  would not guarantee its protection. In contrast, in our setting, individual
  well-behaved activations of the buggy function would be protected.
  \citet{SkorstengaardSTKJFP} note that they support mutually distrustful
  components, but we go a step further. One can think of our hypothetical
  function as ``self-distrustful'' in that the safety of a well-behaved activation
  is not compromised by another, ill-behaved activiation of the same function.

>>>>

That said, our model is meant to apply generically to any enforcement mechanism,
including those that do not take advantage of this separation.
Some mechanisms, like hardware capability systems, might exhibit failstop behavior,
but have no protected policy state---this is represented by a trivial policy
that failstops when the hardware faults. Others might never even failstop, such as
code rewriting techniques that prevent stack safety violations directly, in which
case their policy state does nothing at all.

@@

We now define our machine and enforcement model precisely.

@@

\paragraphx{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of stack safety
is that of \citet{SkorstengaardSTKJFP}. Their
StkTokens system enforces stack safety on hardware equipped with
memory capabilities, similar to those used in the Cheri system~\citep{Woodruff+14,Chisnall+15}, but
with additional support for linear capabilities.
Stack safety is defined as the conjuncton of local state encapsulation (LSE) and
well-bracketed control flow (WBCF).
Informally, LSE means that the current activiation can only access its own frame,
and WBCF means that the topmost frame can only return to
the immediately adjacent frame below. Formally, these properties are defined
with reference to a capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine, an approach termed ``fully abstract overlay semantics.''
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
\begin{enumerate}[label=(\roman*)]
\item Our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
  \citet{SkorstengaardSTKJFP} describe a class of ``unreasonable'' components
  that are not protected, because they are capable of dynamic behavior that can compromise
  their own security. Unreasonable code can be viewed as adversarial, or merely as
  buggy. One can imagine a function that, when called with the wrong
  inputs, behaves unreasonably, but under most circumstances can be
  considered reasonable. \citet{SkorstengaardSTKJFP} would not consider this a
  {\em reasonable component} (as defined in their section 4.3.3), as a reasonable
  component is always reasonable, and so they
  would not guarantee its protection. In our setting, individual
  well-behaved activations of the buggy function would be protected.
  \citet{SkorstengaardSTKJFP} support mutually distrustful
  components; one can think of our hypothetical
  function as ``self-distrustful'' in that the safety of a well-behaved activation
  is not compromised by another, ill-behaved activiation of the same function.
